---
title: "PartB"
author: "Group A4"
date: "06/11/2019"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dataset: Extended Energy Balances (OECD countries) 
Source: Energy Balances of OECD Countries, www.iea.org/statistics

1. Clear the environment and load necessary libraries
```{r preparation, message=FALSE}
rm(list=ls())
library(dplyr)
library(tidyr)
library(readr)
library(readxl)
```

2. Create the reading file loop
The are two cycles in the loop function: The first one is to pick up reference information including country name, flow and unit. The second one copies and combines the content of all files into one data frame $copy.all.contents$
```{r import data, results="hide", message=FALSE}
setwd( "C:/Users/buith/Desktop/Data Management/Group assignment/") #setting working direction 
read.all.file <- list.files("../Group assignment/group_assignment/", recursive = TRUE)
# The first file in the list is a macro xlms file
copy.all.contents <- data.frame()

for (f in  2:length(read.all.file)) {
  this.file <- read.all.file[f]
  this.file.path <- paste0("../Group assignment/group_assignment/", read.all.file[f])
  
  # Pick up contry name, flow and unit
  this.file.header <- read_xlsx(this.file.path, skip = 1, n_max = 3)
  this.file.country <- this.file.header[3,3]
  this.file.flow <- this.file.header[1,3]
  this.file.unit <- this.file.header[2,3]
  
  # Copy all content and merge to one file
  this.file.content <- read_xlsx(this.file.path, skip = 5, na = "..")
  this.file.content.omit <- head(this.file.content, -3) #remove the last 3 rows in each file
  # Add Country, Flow and Unit columns
  this.file.content.omit$Country <-unlist(rep(this.file.country, nrow(this.file.content.omit)))
  this.file.content.omit$Flow <- unlist(rep(this.file.flow, nrow(this.file.content.omit)))
  this.file.content.omit$Unit <- unlist(rep(this.file.unit, nrow(this.file.content.omit)))
 
  copy.all.contents <- rbind(copy.all.contents, this.file.content.omit)
}
```

3. The following codes are used to tidy up the file's structure.
```{r cleanup}
names(copy.all.contents)[1] <- "Year"
# Remove all rows and column that do not include actual values.
copy.all.contents <- copy.all.contents[-which(copy.all.contents$Year == "Time"), ]
copy.all.contents$...2 <- NULL
```

4. The next step is to use gather() function to transpose products from columns to rows.
The final file is $long.copy.all.contents$

```{r transpose}
long.copy.all.contents <- copy.all.contents %>% gather(Product, Value, `Additives/blending components`:Wind)
```

5. Print number of records
```{r}
# The total number of record in the final data frame:
print(nrow(long.copy.all.contents))
head(long.copy.all.contents)
```
```{r}
# The number records for each product: (NA values included)
long.copy.all.contents %>% group_by(Product) %>% summarise(count=n())
# The number records for each product: (NA values excluded)
long.copy.all.contents %>% filter(!is.na(Value)) %>% group_by(Product) %>% summarise(count=n())
```


